\sysfull

### 2017-11-17 paper/coalescing.tex
coalescing: more general strategy definitions

\todo{clarify}{AC}
With this strategy \sys calculates only one parameter $C=C_\text{up}=C_\text{down}$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period. \sys begins execution in a coalesced task with $S_t=0$. When $S_t=C_{t-1}-1$ coalesced tasks have been executed without power interrupt (where $C_{t-1}$ defines $C$ until $t-1$ task commit) a task commit follows. After each commit, \sys assigns new coalescing factor as $C_{t}=\lceil \sum_{k=1}^{t}S_{k-1}/2\rceil$. This implies that if more tasks complete in an operating period, \sys attempts to coalesce more tasks. However, tasks coalescing are increased by a half the speed of the successful task completion. This way we adapt to the coalescing speed/wasted effort trade-off: at the potential power interrupt \sys does not immediately reduce coalescing factor to one, and with successful page commit of coalescing task, \sys gradually increases coalescing factor. In both cases \sys reduces the effect of outliers, i.e. short bursts of energy or short bursts of frequent power interrupts.

%\textbf{Task Coalescing Illustration.} Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks executing in \sys. At left, there is no coalescing and tasks incur the cost of commit at each transition. The center shows the same tasks executing as coalesced groups of two tasks $C=2$, avoiding one commit. At right, \sys coalesces three tasks ($C=3$) avoiding two commits. Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value of $C$ based on $S$. \sys executes tasks in the first operating period, with $C$ equal to half the number of tasks completed in the prior execution period. In the example, $C$ starts at eight. After completing eight tasks, \sys commits and reduces $C$ by half to four. \sys completes two of four tasks and power fails.  During the operating period, \sys executed a total of ten tasks (i.e., $S=10$) and when \sys restarts, $C$ is assigned to five (i.e., half of $S$). \sys completes the five coalesced tasks $S=5$ and after committing, assigns $C$ to half of $S$, which is three (after rounding up). \sys coalesces three tasks and again commits.  At this point, $S=8$; \sys assigns $C=4$ and execution continues.

@@

The history-aware coalescing policy used the number of tasks in the group as a
rough proxy for the energy cost of the task group.
%
With some up-front investment of developer effort and modest time and space
overhead, a more precise proxy for task size can be obtained by profiling the
tasks at design time.
%
The TSHC policy approximates task energy cost from a profile of task execution
times.
%
Profiling execution time instead of direct energy consumption is
advantageous, because the former does not require specialized
current\hyp{}measurement hardware and can be readily done on any
off-the-shelf microcontroller.
%
To use the TSHC policy, the application is first profiled, as will be described
in the experiment below, and the task execution time for each task is written
into a read-only table in non-volatile memory.
%
At runtime, the decision to coalesce or commit at each task boundary proceeds
similarly with the HC policy, but the task completion counter, $S_t$ has units
of ``task execution time''.\todo{be precise}{AC}

To quantify the importance of task size in the coalescing decision, we measured
the distribution of task sizes in our benchmarks (described in
Sections~\ref{sec:results_software}).
%
Each benchmark was run for 10 seconds on non-intermittent power, and a timestamp
in clock cycles was taken at each task boundary. For each execution of
a task, its size was estimated as the maximum difference in timestamps at its
boundaries.
%
The CDF for the distribution of task sizes in Figure~\ref{fig:task_profiling}
shows the wide range of task sizes across and within benchmarks, and explains
the advantage of the TSHC policy.

>>>>

Assuming tasks of similar energy costs, an accessible predictor of a
suitable dynamic task size is the number of successfully completed static tasks
in the recent past.
%
The HC policy keeps track of the number of static task boundaries crossed since
the last boot to set a \emph{target} dynamic task size.
%
This target is expressed in units of ``one static task'' and persisted in
non-volatile memory.
%
When target number of boundaries is crossed, the policy stops coalescing,
thereby setting the size of the dynamic task at the target.

The target is conservatively initialized to zero before the origin task runs
for the first time, i.e. $T_0 \gets 0$.
%
After each successful end of a dynamic task, the policy adjusts the target
upwards, to explore the possibility for larger dynamic tasks.
%
The optimistic upward adjustment of the target is counteracted by the downward
adjustment that takes place after a dynamic task fails to complete after energy
is exhausted, i.e. on each reboot.
%
The adjustments are geometric by a parametrizable factor $\alpha$, i.e.
$T_{i+1} \gets \lceil \alpha T_i \rceil$, where $\alpha  >
1$ for upward and $\alpha < 1$ for downward adjustment.
%
In our experiments, we empirically tuned the controller parameters to $\alpha_\text{up} =
1.5$ and $\alpha_\text{down} = 0.5$.
%
The absolute values of the parameters provide a reaction time that is large
enough to filter out outliers, due to short peaks or lulls in energy supply or
due to disparate task energy cost.
%
The asymmetry reflects conservatism: fast reaction to a decreased energy
availability -- which prioritizes reducing wasted executions over attempts to
extract maximum gain -- and slower reaction to increased energy -- which
cautiously pushes for a larger gain.

%With this strategy \sys calculates only one parameter $C=C_\text{up}=C_\text{down}$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period. \sys begins execution in a coalesced task with $S_t=0$. When $S_t=C_{t-1}-1$ coalesced tasks have been executed without power interrupt (where $C_{t-1}$ defines $C$ until $t-1$ task commit) a task commit follows. After each commit, \sys assigns new coalescing factor as $C_{t}=\lceil \sum_{k=1}^{t}S_{k-1}/2\rceil$. This implies that if more tasks complete in an operating period, \sys attempts to coalesce more tasks. However, tasks coalescing are increased by a half the speed of the successful task completion. This way we adapt to the coalescing speed/wasted effort trade-off: at the potential power interrupt \sys does not immediately reduce coalescing factor to one, and with successful page commit of coalescing task, \sys gradually increases coalescing factor. In both cases \sys reduces the effect of outliers, i.e. short bursts of energy or short bursts of frequent power interrupts.
%
%\textbf{Task Coalescing Illustration.} Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks executing in \sys. At left, there is no coalescing and tasks incur the cost of commit at each transition. The center shows the same tasks executing as coalesced groups of two tasks $C=2$, avoiding one commit. At right, \sys coalesces three tasks ($C=3$) avoiding two commits. Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value of $C$ based on $S$. \sys executes tasks in the first operating period, with $C$ equal to half the number of tasks completed in the prior execution period. In the example, $C$ starts at eight. After completing eight tasks, \sys commits and reduces $C$ by half to four. \sys completes two of four tasks and power fails.  During the operating period, \sys executed a total of ten tasks (i.e., $S=10$) and when \sys restarts, $C$ is assigned to five (i.e., half of $S$). \sys completes the five coalesced tasks $S=5$ and after committing, assigns $C$ to half of $S$, which is three (after rounding up). \sys coalesces three tasks and again commits.  At this point, $S=8$; \sys assigns $C=4$ and execution continues.

@@

The HC policy used the number of tasks in the group as a
rough proxy for the energy cost of the task group.
%
With some up-front investment of developer effort and modest time and space
overhead, a more precise proxy for task size can be obtained by profiling the
tasks at design time.
%
The TSHC policy approximates task energy cost from a profile of task execution
times.
%
Profiling execution time instead of direct energy consumption is advantageous,
because the former does not require any external current\hyp{}measurement
hardware and can be readily done on any off-the-shelf microcontroller.
%
To use the TSHC policy, the application is first profiled, as will be described
in the experiment below, and the task execution time for each task is written
into a read-only table in non-volatile memory.
%
At runtime, the decision for whether to continue coalescing at each task
boundary is determined based on the adaptive target size of the dynamic task as
in the HC policy, but the target and the crossed boundary counter have units of
``task execution time.''

To motivate accounting for task size in coalescing decisions, we measured
the distribution of task sizes in our benchmarks (described in
Sections~\ref{sec:results_software}).
%
Each benchmark was run for 10 seconds on non-intermittent power, and a timestamp
in clock cycles was taken at each task boundary. For each execution of
a task, its size was estimated as the maximum difference in timestamps at its
boundaries.
%
The CDF for the distribution of task sizes in Figure~\ref{fig:task_profiling}
shows the wide range of task sizes across and within benchmarks, and explains
the advantage of the TSHC policy.
